---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fuzzystring

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fuzzystring)](https://CRAN.R-project.org/package=fuzzystring)
[![R-CMD-check](https://github.com/PaulESantos/fuzzystring/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PaulESantos/fuzzystring/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/PaulESantos/fuzzystring/graph/badge.svg)](https://app.codecov.io/gh/PaulESantos/fuzzystring)
<!-- badges: end -->

The goal of fuzzystring is to ...

## Installation

``` r
# Development version from GitHub
# Using pak (recommended)
pak::pak("PaulESantos/fuzzystring")

# Or using remotes
remotes::install_github("PaulESantos/fuzzystring")
```

### Example of `fuzzystring_inner_join`: Correcting misspellings against a dictionary

Often you find yourself with a set of words that you want to combine with a "dictionary"- it could be a literal dictionary (as in this case) or a domain-specific category system. But you want to allow for small differences in spelling or punctuation.

The fuzzyjoin package comes with a set of common misspellings ([from Wikipedia](https://en.wikipedia.org/wiki/Wikipedia:Lists_of_common_misspellings/For_machines)):

```{r}
library(tidyverse)
library(fuzzystring)
data(misspellings)

misspellings
```

```{r words}
# use the dictionary of words from the qdapDictionaries package,
# which is based on the Nettalk corpus.
library(qdapDictionaries)
words <- tbl_df(DICTIONARY)

words
```

As an example, we'll pick 1000 of these words (you could try it on all of them though), and use `fuzzystring_inner_join` to join them against our dictionary.

```{r sub_misspellings}
set.seed(2016)
sub_misspellings <- misspellings %>%
  sample_n(1000)
```

```{r joined, dependson = c("words", "sub_misspellings")}
joined <- sub_misspellings %>%
  fuzzystring_inner_join(words, by = c(misspelling = "word"), max_dist = 1)
```

By default, `fuzzystring_inner_join` uses optimal string alignment (Damerau–Levenshtein distance), and we're setting a maximum distance of 1 for a join. Notice that they've been joined in cases where `misspelling` is close to (but not equal to) `word`:

```{r dependson = "joined"}
joined
```

#### Classification accuracy

Note that there are some redundancies; words that could be multiple items in the dictionary. These end up with one row per "guess" in the output. How many words did we classify?

```{r dependson = "joined"}
joined %>%
  count(misspelling, correct)
```

So we found a match in the dictionary for about half of the misspellings. In how many of the ones we classified did we get at least one of our guesses right?

```{r dependson = "joined"}
which_correct <- joined %>%
  group_by(misspelling, correct) %>%
  summarize(guesses = n(), one_correct = any(correct == word))

which_correct

# percentage of guesses getting at least one right
mean(which_correct$one_correct)

# number uniquely correct (out of the original 1000)
sum(which_correct$guesses == 1 & which_correct$one_correct)
```

Not bad.

#### Reporting distance in the joined output

If you wanted to include the distance as a column in your output, you can use the `distance_col` argument. For example, we may be interested in how many words were *two* letters apart.

```{r joined_dists, dependson = "sub_misspellings"}
joined_dists <- sub_misspellings %>%
  fuzzystring_inner_join(words, by = c(misspelling = "word"), max_dist = 2,
                        distance_col = "distance")

joined_dists
```

Note the extra `distance` column, which in this case will always be less than or equal to 2. We could then pick the closest match for each, and examine how many of our closest matches were 1 or 2 away:

```{r, dependson = "joined_dists"}
closest <- joined_dists %>%
  group_by(misspelling) %>%
  top_n(1, desc(distance)) %>%
  ungroup()

closest

closest %>%
  count(distance)
```

#### Other joining functions

Note that `fuzzystring_inner_join` is not the only function we can use. If we're interested in including the words that we *couldn't* classify, we could have use `fuzzystring_left_join`:

```{r left_joined, dependson = "misspellings"}
left_joined <- sub_misspellings %>%
  fuzzystring_left_join(words, by = c(misspelling = "word"), max_dist = 1)

left_joined

left_joined %>%
  filter(is.na(word))
```

(To get *just* the ones without matches immediately, we could have used `fuzzystring_anti_join`). If we increase our distance threshold, we'll increase the fraction with a correct guess, but also get more false positive guesses:

```{r left_joined2, dependson = "misspellings"}
left_joined2 <- sub_misspellings %>%
  fuzzystring_left_join(words, by = c(misspelling = "word"), max_dist = 2)

left_joined2

left_joined2 %>%
  filter(is.na(word))
```

Most of the missing words here simply aren't in our dictionary.

You can try other distance thresholds, other dictionaries, and other distance metrics (see [stringdist-metrics] for more). This function is especially useful on a domain-specific dataset, such as free-form survey input that is likely to be close to one of a handful of responses.
